---
title: "Kickstarter Data"
output: html_notebook
---

https://www.kaggle.com/kemical/kickstarter-projects 

# Questions

1. Which project categories have viewers that give the highest average donation?
2. Per category, what's the success to failure rate/ratio of all projects?
3. What Kickstarter categories have been most popular over time?

```{r Import Libraries message='hide'}
library(tidyverse)
library(lubridate)
library(scales)
library(MASS)
library(viridis)
library(gridExtra)
```

```{r Import Data message='hide'}

# Use progress=FALSE to remove the progress bar!

ks2018 <- read_csv("./data/ks-2018.csv",col_names=TRUE, progress=FALSE) 
```

```{r Data Cleaning}
# Get rid of spaces in column names, replace with underscore
colnames(ks2018) <- gsub(" ", "_",colnames(ks2018))
```

## Has kickstarter peaked in its popularity as a crowdfunding platform? 
> There was a peak in Kickstarter popularity in 2015, and its popularity has declined since then

```{r Scatterplot}

# Define density function
get_density <- function(x, y, n = 100) {
  dens <- MASS::kde2d(x = x, y = y, n = n)
  ix <- findInterval(x, dens$x)
  iy <- findInterval(y, dens$y)
  ii <- cbind(ix, iy)
  return(dens$z[ii])
}

# Frequency of projects y axis, x axis over time
popOverTime <- ks2018 %>%
    group_by(deadline) %>%
    summarize(count=n()) %>%
    mutate(density=get_density(as.integer(deadline), count))

# Scatterplot with a trend line
popOverTimePlot <- ggplot(popOverTime, aes(x=deadline,y=count,color=density)) +
    geom_point() +
    geom_smooth(method="loess", color="Dark Blue") +
    scale_y_continuous(limits=c(0, 400)) + # Remove values over 500
    scale_x_date(date_breaks = "1 year",
                 labels=date_format("%Y"),
                 limits = as.Date(c('2009-01-01','2018-01-01'))) +
    labs(title="Frequency of Kickstarter Projects over Time") +
    scale_color_viridis() +
    theme(legend.position="none")

print(popOverTimePlot)
```

## Which project categories have viewers that give the highest average donation?
> Technology

```{r Highest Average Donation Per Project Category}

# Roll up dataset into mean pledge per backer per category
pledgeByCat <- ks2018 %>%
    group_by(main_category) %>%
    summarize(
        pledged=sum(usd_pledged, na.rm = TRUE),
        backers=sum(backers, na.rm = TRUE),
        avg=pledged/backers) %>%
    arrange(desc(avg))

# Change main_category into a factor so ggplot will honor the sort order
pledgeByCat$main_category <- factor(pledgeByCat$main_category, levels=pledgeByCat$main_category)

# Plot bar chart of mean pledges
pledgeByCatPlot <- ggplot(pledgeByCat, aes(x=main_category,y=avg,fill=avg)) + geom_col() + labs(title="Avg Pledge By Category Per Backer", x="Category",y="Avg Pledge Per Backer") + theme(axis.text.x=element_text(angle=90),legend.position = "null")

print(pledgeByCatPlot)

```


```{r Success/Failure Ratio of All Projects, Per Category}

successPerCat <- ks2018 %>%
  filter(state %in% c("successful", "failed")) %>%
  group_by(main_category, state) %>%
  summarize(category_count=n()) %>%
  mutate(category_percent=category_count/sum(category_count)) %>%
  arrange(desc(state), category_percent)

# Change main_category into a factor so ggplot will honor the sort order
successPerCat$main_category <- factor(successPerCat$main_category,
                                      levels=unique(successPerCat$main_category))

print(successPerCat$main_category)

successTotals <- ggplot(successPerCat, aes(x=main_category, y=category_percent, fill=state)) + geom_col()
successAndFails <- successTotals + 
  geom_col(position="fill") + 
  scale_fill_discrete(breaks=c("successful", "failed")) + 
  scale_y_continuous(labels=scales::percent) +
  theme(axis.text.x=element_text(angle=90),legend.position = "null")

print(successAndFails)
```

## Trends in kickstarter categories over time

> Hypothesis: kickstarter began with mainly tech projects and then expanded over time

## Trying a heatmap to visualize the trends in categories

> According to our heat map, art projects like film/video and music were most popular. Games and technology did not become popular categories until later.

```{r Heatmap}

# Redo rolling up dataset to include deadline vector
catHeat <- ks2018 %>%
    mutate(year=year(deadline)) %>%
    group_by(main_category,year) %>%
    summarize(
        cat_total=n(), 
        deadline=deadline[1]) %>%
    filter(year < 2018) %>%
    arrange(year)

# Heatmap of categories over time
catHeatPlot <- ggplot(
    catHeat, 
    aes(x=year,y=main_category)) + 
    geom_tile(aes(fill=cat_total)) + 
    scale_fill_gradient(low = 'white', high = 'red') +
    scale_x_continuous(breaks = function(x) unique(floor(pretty(x)))) + 
    labs(title="Most Popular Kickstarter Categories Per Year", x="Year",y="Main Category")

print(catHeatPlot)

```

# Two previous attempts to visualize the above data
> Neither were super readable so I went with the heatmap

## Notes to self re: dplyr
> summarize will strip out anything that isnt in the group by, mutation, or in the summarize itself
> so you need to reference rows in the summarize if you want to include them but they're not in group by or mutate

```{r Line Plot and Density Plot}

# Roll up dataset to number of projects per category per year
catOverTime <- ks2018 %>%
    mutate(year=year(deadline)) %>%
    group_by(main_category,year) %>%
    summarize(cat_total=n()) %>%
    filter(year < 2018) %>%
    arrange(cat_total)

# Line chart of projects per category per year
catOverTimePlot <- ggplot(
    catOverTime, 
    aes(x=year,y=cat_total,colour=main_category)) + geom_line()

#Trying a density plot
# Roll up dataset to number of projects per category per year
# Remove summarize because that frontloads the process that the density plot wants to do
# which is counting the number of events
catOverTime2 <- ks2018 %>%
    mutate(year=year(deadline)) %>%
    group_by(main_category,deadline) %>%
    filter(year < 2018) %>%
    arrange(deadline)

# Density plot of projects
# histogram - plotting frequency of an event in the y axis, versus the scale/groups in the x axis
catDensity <- ggplot(
    catOverTime2,
    aes(x=deadline,fill=main_category,colour=main_category)) + geom_density(alpha = 0.1)

grid.arrange(catOverTimePlot,catDensity, nrow=2)
```




