---
title: "Kickstarter Data"
output: html_notebook
---

https://www.kaggle.com/kemical/kickstarter-projects 

# Questions

1. Which project categories have viewers that give the highest average donation?
2. Per category, what's the success to failure rate/ratio of all projects?
3. What Kickstarter categories have been most popular over time?

```{r Import Libraries}
library(tidyverse)
library(lubridate)
```
```{r Import Data}
ks2018 <- read_csv("./data/ks-2018.csv",col_names=TRUE)
```

```{r Data Cleaning}
# Get rid of spaces in column names, replace with underscore
colnames(ks2018) <- gsub(" ", "_",colnames(ks2018))
```

## Which project categories have viewers that give the highest average donation?
> Technology

```{r Highest Average Donation Per Project Category}

# Roll up dataset into mean pledge per backer per category
pledgeByCat <- ks2018 %>%
    group_by(main_category) %>%
    summarize(
        pledged=sum(usd_pledged, na.rm = TRUE),
        backers=sum(backers, na.rm = TRUE),
        avg=pledged/backers) %>%
    arrange(desc(avg))

# Change main_category into a factor so ggplot will honor the sort order
pledgeByCat$main_category <- factor(pledgeByCat$main_category, levels=pledgeByCat$main_category)

# Plot bar chart of mean pledges
pledgeByCatPlot <- ggplot(pledgeByCat, aes(x=main_category,y=avg,fill=avg)) + geom_col() + labs(title="Avg Pledge By Category Per Backer", x="Category",y="Avg Pledge Per Backer") + theme(axis.text.x=element_text(angle=90),legend.position = "null")

print(pledgeByCatPlot)

```

## Per category, whats the success to failure rate/ratio of all projects?
> Dance has the highest success rate, but a low number of projects overall

```{r Success/Failure Ratio of All Projects, Per Category}

# Roll up dataset
successPerCat <- ks2018 %>%
    group_by(main_category) %>%
    summarize(
        fails = sum(state == "failed", na.rm = TRUE),
        success = sum(state == "successful", na.rm = TRUE),
        success_rate = success / (success + fails) * 100
    ) %>%
    arrange(desc(success_rate))

# Change main_category into a factor so ggplot will honor the sort order
successPerCat$main_category <- factor(successPerCat$main_category, levels=successPerCat$main_category)

# Plot bar chart
successPerCatPlot <- ggplot(successPerCat, aes(x=main_category,y=success_rate,fill=success_rate)) + geom_col() + labs(title="Percent Success Per Category", x="Category",y="Percentage Successes") + theme(axis.text.x=element_text(angle=90),legend.position = "null")

print(successPerCatPlot)

failTotals <- ggplot(successPerCat, aes(x=main_category,y=fails)) + geom_col()

successAndFails <- failTotals + geom_col(position="dodge", aes(y=success,fill=factor(success))) + theme(axis.text.x=element_text(angle=90),legend.position = "null")

print(successAndFails)

```

```{r Success/Failure Ratio of All Projects, Per Category - scala ver}

successPerCat <- ks2018 %>%
  filter(state %in% c("successful", "failed")) %>%
  group_by(main_category, state) %>%
  summarize(category_count=n()) %>%
  mutate(category_percent=category_count/sum(category_count)) %>%
  arrange(desc(state), category_percent)

# Change main_category into a factor so ggplot will honor the sort order
successPerCat$main_category <- factor(successPerCat$main_category,
                                      levels=unique(successPerCat$main_category))

print(successPerCat$main_category)

successTotals <- ggplot(successPerCat, aes(x=main_category, y=category_percent, fill=state)) + geom_col()
successAndFails <- successTotals + 
  geom_col(position="fill") + 
  scale_fill_discrete(breaks=c("successful", "failed")) + 
  scale_y_continuous(labels=scales::percent) +
  theme(axis.text.x=element_text(angle=90),legend.position = "null")

print(successAndFails)
```

## Trends in kickstarter categories over time

> Hypothesis: kickstarter began with mainly tech projects and then expanded over time

```{r Line Plot}

# Roll up dataset to number of projects per category per year
catOverTime <- ks2018 %>%
    select(deadline,main_category,category) %>%
    mutate(year=year(deadline)) %>%
    group_by(main_category,year) %>%
    summarize(cat_total=n()) %>%
    filter(year < 2018) %>%
    arrange(cat_total)

# Line chart of projects per category per year
catOverTimePlot <- ggplot(
    catOverTime, 
    aes(x=year,y=cat_total,colour=main_category)) + geom_line()

print(catOverTimePlot)

# Unfortunately this isn't super readable?
```
## Retrying the above data with a modified dplyr filter and a density plot, to try and get a better visual of the data

> Is there a way to reduce the number of distinct categories so we can only view the top 10% most popular categories?

```{r Density Map}

# Roll up dataset to number of projects per category per year
# Remove summarize because that frontloads the process that the density plot wants to do
# which is counting the number of events
catOverTime2 <- ks2018 %>%
    select(deadline,main_category,category) %>%
    mutate(year=year(deadline)) %>%
    group_by(main_category,deadline) %>%
    filter(year < 2018) %>%
    arrange(deadline)

# Density plot of projects
# histogram - plotting frequency of an event in the y axis, versus the scale/groups in the x axis
catDensity <- ggplot(
    catOverTime2,
    aes(x=deadline,fill=main_category,colour=main_category)) + geom_density(alpha = 0.1)

print(catDensity)
```

## Trying a heatmap to visualize the trends in categories

> According to our heat map, art projects like film/video and music were most popular. Games and technology did not become popular categories until later.

```{r Heatmap}

# Redo rolling up dataset to include deadline vector
catHeat <- ks2018 %>%
    select(deadline,main_category,category) %>%
    mutate(year=year(deadline)) %>%
    group_by(main_category,year) %>%
    summarize(
        cat_total=n(), 
        deadline=deadline[1]) %>%
    filter(year < 2018) %>%
    arrange(year)

# Heatmap of categories over time
catHeatPlot <- ggplot(
    catHeat, 
    aes(x=year,y=main_category)) + 
    geom_tile(aes(fill=cat_total)) + 
    scale_fill_gradient(low = 'white', high = 'red') +
    scale_x_continuous(breaks = function(x) unique(floor(pretty(x)))) + 
    labs(title="Most Popular Kickstarter Categories Per Year", x="Year",y="Main Category")

print(catHeatPlot)

```

## Notes
> summarize will strip out anything that isnt in the group by, mutation, or in the summarize itself
> so you need to reference rows in the summarize if you want to include them but they're not in group by or mutate
> Has kickstarter peaked in its popularity as a crowdfunding platform?